Distributed Topology Discovery in Self-Assembled Nano Networks

### 2014-02-18 PAPER/simulation.tex
Fixes

In order to quantitatively and qualitatively evaluate the proposed approach a
specific simulation environment has been developed, resulting in
the open source and freely available project called
Nanoxim~\cite{nanoxim}.
Nanoxim is a SystemC tool based on a almost rewritten
version of the Noxim Network-on-Chip simulator~\cite{noxim}. While some
complex features have been removed (e.g. wormhole, congestion/topology
aware routing and selection strategies) new features specifically
taylored for the nanoscale scenario were introduced, e.g. the ability to simulate a random
network, the implementation of the \disr{} to obtain the segment topology
and the support for defective links and nodes.

%------------------------------------------------------------------------------
\subsection{Experimental setup}
The following parameters have been taken into account when
performing the \disr{} simulation:
\begin{itemize}
\item {\emph{Size of the network}}: number of nodes, on a range from
10x10 to 100x100 sized networks. 
%\item {\% defective links}: the probability that each link is
%disconnected or not present, from 0 the amount that yelds
%no more segments.
\item {\emph{\% defective nodes}}: the probability that a node is not
working, thus having all its links cannot be utilized during \disr{} setup.  
%Same range as defective links parameter.
\item {\emph{Bootstrap node}}: the node used from upper layer 
to inject the \disr{} process. When not explicitly investigating the
impact of each single bootstrap choice, a set representative regions have been
considered, e.g. the central part of the network and the edge corners.
\end{itemize}
To present the results, the following evaluation metrics have been adopted:
\begin{itemize}
\item{Node coverage}: this is the fraction of nodes that
are assigned to a segment. In the ideal case, all the non defective
nodes should be assigned, so this metric is useful to show how some
disconnected regions can negatively impact on the whole \disr{}
effectiveness.
\item{\emph{Latency}}: this measures how the cycles required to complete the
segment assignment scales for increasing network sizes and defect rates
\item{\emph{Bootstrap node effect}}: this evaluates the impact of the chosen
bootstrap node on the node coverage.
\end{itemize}

%Further, to evaluate how \disr{} compares against the centralized segment
%based algorithm, the following metrics have been adopted:
%\begin{itemize}
%%\item {Average Path Lenght}:
%%\item {Average Link Weight}:
%\item {Unidirectional restrictions:}
%\item {Number of Segments}:
%\end{itemize}

@@

% TODO: long paper
\subsection{Other Optimisations}
Some optimisation parameters, which demostrated to improve the \disr{}
results have been fixed to some good values, but are not subject to
further investigation; once again the focus here is not the
optimal set of segments, but just demostrating a working approach. 
These are optimisations are:
\begin{itemize}
\item \emph{cycle\_links}: max number of retries across the set of links of
each node. While searching for a free link due an incoming request,
the same request is cancelled after a given number of tries. This
gives to preceding node change to test a different paths. Default is set to 1.
\item \emph{bootstrap\_timeout}: number of time units that a bootstrap node
should wait before assuming that a livelock in the starting segment
process has occurred. In the worst case, we can imagine that longest
path required is the one returning to the bootstrap node after having
travelled across all the links. So, althought this is just an extreme
situation, a good upper limit can be safely be set to NxN.
\item \emph{bootstrap\_immunity}: in order to avoid the failure of the whole \disr{}
setup process, a bootstrap node should not have defective links.
Enabling this optimisation, a bootstrap node is immune to defects.
We may think of a pre-bootstrap phase that properly selects (from upper
layer via) a bootstrap node which is tested as properly connected. We
enabled this optimisation, however empirical tests showed us that only
simulations using bootstrap nodes placed on edges would be heavily
affected, since these are nodes starting with a lower number of links,
e.g. corner nodes could only have two connected direction, so even a
defective link could prevent starting segment packet to return to
bootstrap to close the loop a create the segment.
\end{itemize}

>>>>

In order to quantitatively and qualitatively evaluate the proposed approach a
specific simulation environment has been developed, resulting in
the open source and freely available project called
Nanoxim~\cite{nanoxim}.
Nanoxim is a SystemC tool based on a almost rewritten
version of the Noxim Network-on-Chip simulator~\cite{noxim}. While some
complex features have been removed (e.g. wormhole, congestion/topology
aware routing and selection strategies) new features specifically
tailored for the nanoscale scenario were introduced, e.g. the ability to simulate a random
network, the implementation of the \disr{} to obtain the segment topology
and the support for defective links and nodes.

%------------------------------------------------------------------------------
\subsection{Experimental setup}
The following parameters have been taken into account when
performing the \disr{} simulation:
\begin{itemize}
\item {\emph{Size of the network}}: number of nodes, on a range from
10x10 to 100x100 sized networks. 
%\item {\% defective links}: the probability that each link is
%disconnected or not present, from 0 the amount that yields
%no more segments.
\item {\emph{\% defective nodes}}: the probability that a node is not
working, thus having all its links cannot be utilized during \disr{} setup.  
%Same range as defective links parameter.
\item {\emph{Bootstrap node}}: the node used from upper layer 
to inject the \disr{} process. When not explicitly investigating the
impact of each single bootstrap choice, a set representative regions have been
considered, e.g. the central part of the network and the edge corners.
\end{itemize}
To present the results, the following evaluation metrics have been adopted:
\begin{itemize}
\item{\emph{Node coverage}}: this is the fraction of nodes that
are assigned to a segment. In the ideal case, all the non defective
nodes should be assigned, so this metric is useful to show how some
disconnected regions can negatively impact on the whole \disr{}
effectiveness.
\item{\emph{Latency}}: this measures how the cycles required to complete the
segment assignment scales for increasing network sizes and defect rates
\item{\emph{Bootstrap node effect}}: this evaluates the impact of the chosen
bootstrap node on the node coverage.
\end{itemize}

%Further, to evaluate how \disr{} compares against the centralized segment
%based algorithm, the following metrics have been adopted:
%\begin{itemize}
%%\item {Average Path Length}:
%%\item {Average Link Weight}:
%\item {Unidirectional restrictions:}
%\item {Number of Segments}:
%\end{itemize}

@@

% TODO: long paper
\subsection{Other Optimizations}
Some optimization parameters, which demonstrated to improve the \disr{}
results have been fixed to some good values, but are not subject to
further investigation; once again the focus here is not the
optimal set of segments, but just demonstrating a working approach. 
These are optimizations are:
\begin{itemize}
\item \emph{cycle\_links}: max number of retries across the set of links of
each node. While searching for a free link due an incoming segment request,
the same request is cancelled after a given number of tries. This
gives to the preceding node receives a segment cancel and has the
chance of testing a different path instead of waiting indefinitely.
Default is set to 1, i.e. each link is tried one time.
\item \emph{time to live (TTL}: when a segment request is cancelled
along a certain path (using a segment cancel), a TTL field is
decreased in the request packet, so that requests that have been
denied too much times can desist, even if free links for going on are
available. This is not only a optimisation for shortening 
requests processes, but also solves some critical situations in which
request packets looped in a particular path without neither succeeding
nor failing.
\item \emph{bootstrap\_timeout}: number of time units that a bootstrap node
should wait before assuming that a livelock in the starting segment
process has occurred. In the worst case, we can imagine that longest
path required is the one returning to the bootstrap node after having
traveled across all the links. So, although this is just an extreme
situation, a good upper limit can be safely be set to $N \times N$.
\item \emph{bootstrap\_immunity}: in order to avoid the failure of the whole \disr{}
setup process, a bootstrap node should not have defective links.
Enabling this optimization, a bootstrap node is immune to defects.
We may think of a pre-bootstrap phase that properly selects (from upper
layer via) a bootstrap node which is tested as properly connected. We
enabled this optimization, however empirical tests showed us that only
simulations using bootstrap nodes placed on edges would be heavily
affected, since these are nodes starting with a lower number of links,
e.g. corner nodes could only have two connected direction, so even a
defective link could prevent starting segment packet to return to
bootstrap to close the loop a create the segment.
\end{itemize}
